---
title: "Kinetic Model Coefficient Analysis"
format: 
  html:
    math: mathjax
  pdf: default
  docx: default
author: Marc Pérez
date: 2025-05-22
---

```{r setup}
#| include: false
#| message: false
#| warning: false


suppressPackageStartupMessages({
  library(multcomp)
  library(car)
  library(tidyr)
  library(lme4)
  library(ggplot2)
  library(ggtext)
  library(ggpmisc)
  library(nlme)
  library(latex2exp)
  library(kableExtra)
  library(broom)
  library(dplyr)
  library(MuMIn)
})

options(warn = -1)
RES <- readRDS("~/Documents/Master Thesis/Master-Thesis-P-kinetics/data/RES.rds")
D <- RES$D
d <- RES$data
D$annual_P_balance2 <-   D$fert_P_tot-D$harv1_mp_Puptake
```

Model Agroscope
$$Y_{rel}\sim A*(1-e^{rate*P_{CO_2}+Env})$$

Wir ersetzen nur rate mit unserer Schätzung k:
$$Y_{rel}\sim A*(1-e^{k*P_{CO_2}+Env})$$

Sind unsere Modelparameter gute Prediktoren??
$$Y_{rel}\sim A*(1-e^{k*PS+Env} )$$

Es gibt noch die Kovariaten Niederschlag pro Jahr, Jahresdurchschnittstemperatur und Temperatur in Jugendphase



```{r}
#|code-fold: true
#|tidy: true






library(GGally)

ggpairs(D, 
  aes(col=Site, shape = Treatment,alpha = 0.6), 
  columns = c("soil_0_20_P_AAE10", "soil_0_20_P_CO2", "PS", "k", "kPS"),
  lower = list(continuous = wrap("points", size = 1.3)),
  upper = list(continuous = "blank", combo = "blank", discrete = "blank"))  # Adjust size here
  




p6 <- ggplot(D,aes(y=soil_0_20_P_AAE10, x=soil_0_20_P_CO2, col=Site, size = Treatment)) +
  geom_point(shape = 7) + 
  scale_x_log10() + scale_y_log10() +
  labs(x=TeX("$P_{H_2O10}(mg/kg Soil)$"),
         y=TeX("$P_{AAEDTA}(mg/kg Soil)$")); p6
  
p7 <- ggplot(D,aes(y=PS, x=soil_0_20_P_CO2, col=Site, size = Treatment)) +
  geom_point(shape = 7) +
  scale_x_log10() + scale_y_log10() +
  labs(x=TeX("$P_{CO_2}(mg/kg Soil)$"),
          y=TeX("$PS(mg/kg Soil)$")); p7

p8 <- ggplot(D,aes(y=k, x=soil_0_20_P_CO2, col=Site, size = Treatment)) +
  geom_point(shape = 7) +
  scale_x_log10() +
  labs(x=TeX("$P_{CO_2}(mg/kg Soil)$"),
         y=TeX("$k(1/s)$")); p8

p9 <- ggplot(D,aes(y=k*PS, x=soil_0_20_P_CO2, col=Site, size = Treatment)) +
  geom_point(shape = 7) +
  scale_x_log10() + scale_y_log10() +
  labs(x=TeX("$P_{CO_2}(mg/kg Soil)$"),
         y=TeX("$v=k*PS(mg/s*kg Soil)$"));p9


p11 <- ggplot(D,aes(y=PS, x=soil_0_20_P_AAE10, col=Site, size = Treatment)) +
  scale_x_log10() + scale_y_log10() +
  geom_point(shape = 7) +
  labs(x=TeX("$P_{AAEDTA}(mg/kg Soil)$"),
         y=TeX("$PS(mg/kg Soil)$")); p11

p12 <- ggplot(D,aes(y=k, x=soil_0_20_P_AAE10, col=Site, size = Treatment)) +
  geom_point(shape = 7) +
   scale_x_log10() + scale_y_log10() +
  labs(x=TeX("$P_{AAEDTA}(mg/kg Soil)$"),
         y=TeX("$k(1/s)$"))

p12

p13 <- ggplot(D,aes(y=k*PS, x=soil_0_20_P_AAE10, col=Site, size = Treatment)) +
  scale_x_log10() + scale_y_log10() +
  geom_point(shape = 7) +
  labs(x=TeX("$P_{AAEDTA}(mg/kg Soil)$"),
         y=TeX("$log(v)=log(k*PS)(mg/s*kg Soil)$"))

p13




```


Nun noch die Linearen Regressionen, die ausstehend sind:


`(1|year) + (1|Site)  + (1|Site:block) + (Treatment|Site)`

Random intercept per year and site, block nested in site. and Treatment nested in site (could also be modelled as a random slope to allow for correlations)

wir sind abe nicht an einem Treatment effekt interesseiert. drum verwerfen wir Treatment als Random UND Fixed effekt.

1. Vergleiche PS, k und kPS mit 



```{r}
#|code-fold: true
#|tidy: true


# Wovon hängen Modelparameter ab?

library(lmerTest)

fit.soil.PS  <- lmer(log(PS)     ~ soil_0_20_clay+ soil_0_20_pH_H2O + soil_0_20_Corg + soil_0_20_silt + Treatment + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)
# fit.soil.PS2 <- lmer(log(PS)     ~ soil_0_20_clay+ soil_0_20_pH_H2O + soil_0_20_Corg + soil_0_20_silt + Treatment + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)
fit.soil.k   <- lmer(k           ~ soil_0_20_clay+ soil_0_20_pH_H2O + soil_0_20_Corg + soil_0_20_silt + Treatment + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)
fit.soil.kPS <- lmer(I(log(k*PS))~ soil_0_20_clay+ soil_0_20_pH_H2O + soil_0_20_Corg + soil_0_20_silt + Treatment + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)
fit.soil.kPS2<- lmer(I(k*log(PS))~ soil_0_20_clay+ soil_0_20_pH_H2O + soil_0_20_Corg + soil_0_20_silt + Treatment + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)
fit.soil.CO2 <- lmer(log(soil_0_20_P_CO2)~ soil_0_20_clay+ soil_0_20_pH_H2O + soil_0_20_Corg + soil_0_20_silt + Treatment + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)
fit.soil.AAE10<-lmer(log(soil_0_20_P_AAE10)~ soil_0_20_clay+ soil_0_20_pH_H2O + soil_0_20_Corg + soil_0_20_silt + Treatment + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)





r.squaredGLMM(fit.soil.kPS)
r.squaredGLMM(fit.soil.kPS2)

fit.kin.Yrel     <- lmer(Ymain_rel         ~ k * log(PS)  + (1|year) + (1|Site)  + (1|Site:block), D)

fit.kin.Ynorm    <- lmer(Ymain_norm       ~ k + log(PS)  + (1|year) + (1|Site)  + (1|Site:block), D, subset = Treatment != "P166")
r.squaredGLMM(fit.kin.Ynorm)


fit.kin.Ynorm    <- lmer(Ymain_norm       ~ k * log(PS)  + (1|year) + (1|Site)  + (1|Site:block), D, subset = Treatment != "P166")
r.squaredGLMM(fit.kin.Ynorm)

fit.kin.Ynorm    <- lmer(Ymain_norm       ~ k * log(PS)  + (1|year), D, subset = Treatment != "P166")
r.squaredGLMM(fit.kin.Ynorm)


fit.kin.Ynorm    <- lmer(Ymain_norm       ~ k * log(PS)  + (1|year) + (1|Site)  + (1|Site:block), D)
r.squaredGLMM(fit.kin.Ynorm)

fit.kin.Ynorm    <- lmer(Ymain_norm       ~ k * log(PS)  + (1|year) + (1|Site)  + (1|Site:block), D, subset = Treatment == "P0")
r.squaredGLMM(fit.kin.Ynorm)
summary(fit.kin.Ynorm, ddf="Kenward-Roger")

fit.kin.Ynorm    <- lmer(Ymain_norm       ~ k * log(PS)  + (1|year) + (1|Site)  + (1|Site:block), D, subset = Treatment == "P100")
r.squaredGLMM(fit.kin.Ynorm)

fit.kin.Ynorm    <- lmer(Ymain_norm       ~ k * log(PS)  + (1|year) + (1|Site)  + (1|Site:block), D, subset = Treatment == "P166")
r.squaredGLMM(fit.kin.Ynorm)



fit.grud.CO2.Ynorm     <- lmer(Ymain_norm    ~ log(soil_0_20_P_CO2) + (1|year) + (1|Site)  + (1|Site:block), D, subset = Treatment != "P166")
r.squaredGLMM(fit.grud.CO2.Ynorm)
fit.grud.AAE10.Ynorm   <- lmer(Ymain_norm    ~ log(soil_0_20_P_AAE10) + (1|year) + (1|Site)  + (1|Site:block), D, subset = Treatment != "P166")
r.squaredGLMM(fit.grud.AAE10.Ynorm)
fit.grud.CO2.AAE10.Ynorm <- lmer(Ymain_norm  ~ log(soil_0_20_P_CO2) * log(soil_0_20_P_AAE10) + (1|year) + (1|Site)  + (1|Site:block), D, subset = Treatment != "P166")
r.squaredGLMM(fit.grud.CO2.AAE10.Ynorm)
# compare with k*log(PS)
fit.kin.Ynorm   <- lmer(Ymain_norm       ~ k*log(PS) + (1|year) + (1|Site)  + (1|Site:block), D, subset = Treatment != "P166")
r.squaredGLMM(fit.kin.Ynorm)

fit.kin.Ynorm    <- lmer(Ymain_norm       ~ k * log(PS) + (1|year) + (1|Site)  + (1|Site:block), D, subset = Treatment != "P166")
fit.kin.Pexport  <- lmer(harv1_mp_Puptake  ~ k * log(PS) + (1|year) + (1|Site)  + (1|Site:block), D)
fit.kin.Pbalance <- lmer(annual_P_balance2 ~ k * log(PS) + (1|year) + (1|Site)  + (1|Site:block), D)

car::vif(lm(Ymain_norm       ~ (k) * log(PS)  + crop, D))
car::vif(lm(Ymain_norm       ~ scale(k) * scale(log(PS))   + crop, D))
car::vif(lm(Ymain_norm       ~ k + log(PS)   + crop, D))
car::vif(lm(Ymain_norm       ~ I(k * (PS)) + k + log(PS)    + crop, D))

with(D, hist(I(exp(k) * (PS))))
with(D, hist(log(PS)))

r.squaredGLMM(fit.kin.Ynorm)

anova(fit.soil.PS)
#summary(glht(fit.PS))
# Fazit: PS wird von treatment stark beeinfluss, k eher nicht (dafür von site)
anova(fit.soil.k)
summary(glht(fit.soil.k))
anova(fit.soil.kPS)
summary(glht(fit.soil.kPS))
anova(fit.kin.Yrel)
anova(fit.kin.Ynorm)
anova(fit.kin.Pexport)
anova(fit.soil.CO2)
anova(fit.soil.AAE10)
anova(fit.kin.Pbalance)
summary(fit.kin.Pbalance)
fit.kin.Pbalance |> r.squaredGLMM()
fit.kin.Pexport |> r.squaredGLMM()
fit.kin.Yrel |> r.squaredGLMM()
fit.kin.Ynorm |> r.squaredGLMM()
# Verhalten der Modelparameter und Ertragsdaten auf P-CO2 und P-AAE10

```

Since we now model two measurement methods, we do not expect correlations by Site/year/etc
```{r}
#|code-fold: true
#|tidy: true
# fit.PS       <- lm(PS            ~ soil_0_20_P_CO2 + soil_0_20_P_AAE10, D)
fit.grud.PS       <- lm(log(PS)         ~ log(soil_0_20_P_CO2) + log(soil_0_20_P_AAE10), D)
fit.grud.k        <- lm(k               ~ log(soil_0_20_P_CO2) + log(soil_0_20_P_AAE10), D)
fit.grud.kPS      <- lm(I(log(k*PS))    ~ log(soil_0_20_P_CO2) + log(soil_0_20_P_AAE10), D)

fit.grud.CO2.Yrel     <- lmer(Ymain_rel       ~ log(soil_0_20_P_CO2)  + Treatment + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)
fit.grud.AAE10.Yrel     <- lmer(Ymain_rel       ~ log(soil_0_20_P_AAE10) + Treatment + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)
fit.grud.Yrel     <- lmer(Ymain_rel       ~ log(soil_0_20_P_CO2) + log(soil_0_20_P_AAE10) + Treatment + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)


# this is hopeless, since cannot log becaus of 0's
fit.CO2.Pexport     <- lmer(harv1_mp_Puptake     ~ log(soil_0_20_P_CO2) + Treatment + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)

fit.AAE10.Pexport     <- lmer(harv1_mp_Puptake ~ log(soil_0_20_P_AAE10) + Treatment + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)

fit.grud.Pexport     <- lmer(harv1_mp_Puptake     ~ log(soil_0_20_P_CO2) + log(soil_0_20_P_AAE10) + Treatment + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)


fit.CO2.Pbalance     <- lmer(annual_P_balance2      ~ log(soil_0_20_P_CO2)  + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)


fit.AAE10.Pbalance     <- lmer(annual_P_balance2      ~ log(soil_0_20_P_AAE10)  + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)


fit.grud.Pbalance     <- lmer(annual_P_balance2      ~ log(soil_0_20_P_CO2) + log(soil_0_20_P_AAE10)  + (1|year) + (1|Site)  + (1|Site:block) + (1|Site:Treatment), D)



#summary(glht(fit.PS))
# Fazit: PS wird von treatment stark beeinfluss, k eher nicht (dafür von site)
summary(glht(fit.grud.PS))
summary(glht(fit.grud.k))
summary(glht(fit.grud.kPS))
summary(fit.grud.Yrel)
summary(fit.grud.Pexport)
summary(fit.grud.Pbalance)
fit.grud.Pbalance |> r.squaredGLMM()
fit.grud.Pexport |> r.squaredGLMM()
fit.grud.Yrel |> r.squaredGLMM()


save.image(file = "~/Documents/Master Thesis/Master-Thesis-P-kinetics/data/results_coefficient_analysis")

```



```{r}
create_coef_table <- function(lmer_models, sig_threshold = 0.001, covariate_order = NULL) {

  # Extract coefficients and p-values from lmerTest::lmer objects
  extract_coef_info <- function(model) {
    coef_matrix <- summary(model)|> coef()
    estimates <- coef_matrix[, 1]
    p_values <- coef_matrix[, ncol(coef_matrix)]
    # Format with significance stars
    formatted_coef <- sapply(seq_along(estimates), function(i) {
    est_str <- sprintf("%.3f", estimates[i])
    stars <- if (p_values[i] < 0.001) "***" else
    if (p_values[i] < 0.01) "** " else
    if (p_values[i] < 0.05) "*  " else  ""
    paste0(stars, est_str)
    })
    names(formatted_coef) <- rownames(coef_matrix)
    return(formatted_coef)
  }

  # Extract R-squared values
  extract_r_squared <- function(model) {
  r2_values <- r.squaredGLMM(model)
  return(c(
    R2m = sprintf("%.3f", r2_values[1, "R2m"]),
    R2c = sprintf("%.3f", r2_values[1, "R2c"])
  ))
  }

  # Extract coefficients for all models
  all_coefs <- lapply(lmer_models, extract_coef_info)

  # Extract R-squared for all models
  all_r_squared <- lapply(lmer_models, extract_r_squared)

  # Get all unique coefficient names
  all_covariate_names <- unique(unlist(lapply(all_coefs, names)))

  # Apply covariate ordering
  if (is.null(covariate_order)) {
  # Default: Intercept first, then alphabetical
  covariate_order <- c("(Intercept)", sort(all_covariate_names[all_covariate_names != "(Intercept)"]))
  }

  # Filter to only include covariates that exist
  covariate_order <- covariate_order[covariate_order %in% all_covariate_names]

  # Add R-squared rows to the ordering
  final_order <- c(covariate_order, "R2m", "R2c")

  # Create results matrix
  results_matrix <- matrix("",
  nrow = length(final_order),
  ncol = length(lmer_models),
  dimnames = list(final_order, names(lmer_models)))

  # Fill in the coefficients
  for (model_name in names(lmer_models)) {
  model_coefs <- all_coefs[[model_name]]
  for (covar in names(model_coefs)) {
  if (covar %in% covariate_order) {
  results_matrix[covar, model_name] <- model_coefs[covar]
  }
  }
  }

  # Fill in the R-squared values
  for (model_name in names(lmer_models)) {
  r2_values <- all_r_squared[[model_name]]
  results_matrix["R2m", model_name] <- r2_values["R2m"]
  results_matrix["R2c", model_name] <- r2_values["R2c"]
  }

  # Convert to data frame for kable
  results_df <- data.frame(Covariate = rownames(results_matrix),
  results_matrix,
  stringsAsFactors = FALSE)

  results_df
}
```


```{r}
lmer_models <- list(
  PS = fit.soil.PS,
  k = fit.soil.k,
  'log(k*PS)' = fit.soil.kPS,
  'k*log(PS)' = fit.soil.kPS2,
  CO2 = fit.soil.CO2,
  AAE10 = fit.soil.AAE10,
  fail = fit.kin.Yrel
)



coef_table_soil <- create_coef_table(lmer_models, sig_threshold = 0.001)
kable(coef_table_soil,
row.names = FALSE,
align = c("l", rep("r", ncol(coef_table_soil) - 1)),
caption = "Coefficient Table for Soil covariates. 
Significant codes:  0 '\\*\\*\\*' 0.001 '\\*\\*' 0.01 '\\*' 0.05")




lmer_models_yield <- list(
  CO2_Ynorm = fit.grud.CO2.Ynorm,
  AAE10_Ynorm = fit.grud.AAE10.Ynorm,
  Grud_Ynorm = fit.grud.CO2.AAE10.Ynorm,
  Kin_Ynorm = fit.kin.Ynorm,
  CO2_Yrel = fit.grud.CO2.Yrel,
  AAE10_Yrel = fit.grud.AAE10.Yrel,
  Grud_Yrel = fit.grud.Yrel,
  Kin_Yrel = fit.kin.Yrel
)

lmer_models_balance <- list(
  CO2_Pbalance = fit.CO2.Pbalance,
  AAE10_Pbalance = fit.AAE10.Pbalance,
  Grud_Pbalance = fit.grud.Pbalance,
  Kin_Pbalance = fit.kin.Pbalance
)

lmer_models_export <- list(
  CO2_Pexport = fit.CO2.Pexport,
  AAE10_Pexport = fit.AAE10.Pexport,
  Grud_Pexport = fit.grud.Pexport,
  Kin_Pexport = fit.kin.Pexport
)


coef_table_yield <- create_coef_table(lmer_models_yield, sig_threshold = 0.001)
kable(coef_table_yield,
row.names = FALSE,
align = c("l", rep("r", ncol(coef_table_yield) - 1)),
caption = "Coefficient Table for Ynorm and Yrel. 
Significant codes:  0 '\\*\\*\\*' 0.001 '\\*\\*' 0.01 '\\*' 0.05")











save.image(file = "~/Documents/Master Thesis/Master-Thesis-P-kinetics/data/results_coefficient_analysis")


```

